---
title: "Assignment_2"
author: "Abhinav Garg"
date: "1/25/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the Data 
```{r Load}
library(nycflights13) # load library 
data(flights) # load data on flights 
```
## 1. 
##Let's explore flights from NYC to Seattle. Use the flights dataset to answer the following questions

a) How many flights were there to and from NYC in 2013 ? 
```{r 1.a}
f <- flights
nyc_dest <- f[f$dest %in% c('LGA','JFK','EWR'),] # FLights to NYC airport
a<- nrow(f) # all Flights originate from NYC Airports
b<- nrow(nyc_dest)

# Number of flights from and to NYC
a+b
```
b) How many flights were from NYC Airports to Seattle in 2013 ? 
```{r 1.b}
nyc_sea <- f[(f$dest %in% c('SEA') & f$origin %in% c('LGA','EWR','JFK')),] # Origin is one of NYC Airports and dest is SEA
nrow(nyc_sea)
```
c) How many airlines fly from NYC to Seattle? 
```{r 1.c}
nyc_sea <- f[(f$dest %in% c('SEA') & f$origin %in% c('LGA','EWR','JFK')),] # Origin is one of NYC Airports and dest is SEA
nyc_sea_carrier <- unique(nyc_sea$carrier) # Unique carriers that fly from NYC to SEA
length(nyc_sea_carrier) # Number of unique Carriers 
```
d) What is the average arrival delay for flights from NYC to Seattle ?
```{r 1.d}
nyc_sea <- f[(f$dest %in% c('SEA') & f$origin %in% c('LGA','EWR','JFK')),] # Origin is one of NYC Airports and dest is SEA
avgarr_delay <- mean(nyc_sea$arr_delay,na.rm = T)
avgarr_delay
```

##2. 
##Flights are often delayed. Consider the following questions exploring delay patterns.

a) What is the mean arrival delay time? What is the median arrival delay time?
```{r 2.a}
mean_arrdelay <- mean(f$arr_delay, na.rm = T)
# Mean Arrival Delay Time 
mean_arrdelay
median_arrdelay <- median(f$arr_delay, na.rm = T)
# Median Arrival Delay Time
median_arrdelay

```
b) What does a negative arrival delay mean ?
```{r 2.b}
?flights # Gives description about variables 

```
Negative times represents early arrivals 

c) Plot a histogram of arrival delay times. Does the answers you obtained in (a)
consistent with the shape of the delay time distribution?
```{r 2.c}
hist(f$arr_delay,xlim = c(-200,300),breaks = 100,xlab = 'Arrival Delay Time',main= 'Frequency of Arrival Delay Time')
# Yes the answers obtained in (a) are consistent with the shape 
```

d) 