---
title: "Assignment_3"
author: "Abhinav Garg"
date: "2/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1
For each of the following situations, state whether the parameter of interest is a mean or a
proportion. It may be helpful to examine whether individual responses are numerical or
categorical.

a) (1 pt) In a survey, one hundred college students are asked how many hours per week they spend on the Internet.

Ans: These responses are numerical and the parameter of interest is a mean 

b) (1 pt) In a survey, one hundred college students are asked: “What percentage of the time you spend on the Internet is part of your course work?”

Ans: These responses could be numerical or categorical depending on the users response as it is an open ended question. Parameter of interest would be proportion 

c) (1 pt) In a survey, one hundred college students are asked whether or not they cited information from Wikipedia in their papers.

Ans: These responses would be categorical as they would be answered in yes / no / no response. This parameter would be proportion 

d) (1 pt) In a survey, one hundred college students are asked what percentage of their total weekly spending is on alcoholic beverages.

Ans: The parameter of interest here is proportion 

e) (1 pt) In a sample of one hundred recent college graduates, it is found that 85 percent expect to get a job within one year of their graduation date.

Ans: The parameter of interest here is proportion 

## 2

(5 pt) In 2013, the Pew Research Foundation reported that “45% of U.S. adults report that they live with one or more chronic conditions”. However, this value was based on a sample, so it may not be a perfect estimate for the population parameter of interest on its own. The study reported a standard error of about 1.2%, and a normal model may reasonably be used in this setting. Create a 95% confidence interval for the proportion of U.S. adults who live with one or more chronic conditions. Also interpret the confidence interval in the context of the study.

```{r 2}
u <- 45
se <- 1.2
left <- 45 - 1.96*se
right <- 45 + 1.96*se
left
```

```{r 2.a}
right 
```
So as we can see the confidence interval is between (42.648, 47.352)
With 95% confidence we can say that the average number of adults that live with one or more chronic conditions lies between 42.648 and 47.352 


## 3 

The nutrition label on a bag of potato chips says that a one ounce (28 gram) serving of potato chips has 130 calories and contains ten grams of fat, with three grams of saturated fat. A random sample of 35 bags yielded a sample mean of 136 calories with a standard deviation of 17 calories.

a) Write down the null and alternative hypotheses for a two-sided test of whether the nutrition label is lying.

Null Hypothesis : The mean of a one ounce serving of potato chips is 130 calories 

Alternate Hypothesis : Th mean of a one ounce seving of potato chips is greater than 130 calories 

b) (4 pt) Calculate the test statistic and find the p value.

Assuming the null hypothesis to be true : u  = 130 
x_bar = 136 
n = 35
sd = 17 

Z = x_bar - u / ( sd / sqrt(n) )

```{r 3.b}
u  = 130 
x_bar = 136 
n = 35
sd = 17 
num <- x_bar - u 
den <- sd / sqrt(n)
Z <- num /den 
2*pnorm(-abs(Z))
```

(2 pt) If you were the potato chip company would you rather have your alpha = 0.05 or
0.025 in this case? Why?

since the Z-score is 2.088 which is approximately 2 std deviations away from the mean. If the significance level is 0.05 which corresponds to 2 standard deviations from the mean, we would have to reject the null hypothesis that the calories in one serving is 130 calories and accept that it is more than 130 calories. However if the significance level is 0.025 we would be within the confidence interval and accept the null hypothesis.But that being said, we cannot assume the significance level after calculating or taking the samples as this is ethically not correct. 


## 4 

Regression was originally used by Francis Galton to study the relationship between parents and children. He wondered if he could predict a man’s height based on the height of his father? This is the question we will explore in this problem. You can obtain data similar to that used by Galton as follows:

```{r 4}
library(UsingR)
height <- get("father.son")

```

a) (5 pt) Perform an exploratory analysis of the father and son heights. What does the relationship look like? Would a linear model be appropriate here?
```{r 4.a}
plot(sheight ~ fheight, data = height, xlab = "Father's height", ylab="Son's height" )
title("Plot of Father-Son Height")
cor(height$sheight, height$fheight)
```
We can see that there is a positive correlation between the son and father's height, therefore a linear model could be utilized here. 


b) (5 pt) Use the lm function in R to fit a simple linear regression model to predict son’s
height as a function of father’s height. Write down the model,

ysheight = β0 + β1 x fheight

filling in estimated coefficient values and interpret the coefficient estimates.
```{r 4.b}
mod3 <- lm(height$sheight ~ height$fheight)
summary(mod3)
```

c) (5 pt) Find the 95% confidence intervals for the estimates. You may find the confint() command useful.
```{r 4.c}
confint(mod3,level=0.95)
```

d) (5 pt) Produce a visualization of the data and the least squares regression line.

```{r 4.d}
library(ggplot2)
g<- ggplot(data = height,mapping = aes(x= fheight, y = sheight))
g <- g + geom_point() + xlab("Father's Height") + ylab("Son's Heights")
g <- g + labs(title = "Plot of Father- Son Height")
g <- g + geom_smooth(method = 'lm', se = TRUE)
g
```


## 5 



An investigator is interested in understanding the relationship, if any, between the analytical skills of young gifted children and the father’s IQ, the mother’s IQ, and hours of educational TV. The data are here:

```{r 5}
library(openintro)
data(gifted)
```

```{r 5.a}
mod1 <- lm(fatheriq ~ score, data = gifted)
summary(mod1)
```

```{r 5.b}
mod2 <- lm(motheriq ~ score, data = gifted)
summary(mod2)
```