---
title: "Assignment_3"
author: "Abhinav Garg"
date: "2/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1
For each of the following situations, state whether the parameter of interest is a mean or a
proportion. It may be helpful to examine whether individual responses are numerical or
categorical.

a) (1 pt) In a survey, one hundred college students are asked how many hours per week they spend on the Internet.

Ans: These responses are numerical and the parameter of interest is a mean 

b) (1 pt) In a survey, one hundred college students are asked: “What percentage of the time you spend on the Internet is part of your course work?”

Ans: These responses could be numerical or categorical depending on the users response as it is an open ended question. Parameter of interest would be proportion 

c) (1 pt) In a survey, one hundred college students are asked whether or not they cited information from Wikipedia in their papers.

Ans: These responses would be categorical as they would be answered in yes / no / no response. This parameter would be proportion 

d) (1 pt) In a survey, one hundred college students are asked what percentage of their total weekly spending is on alcoholic beverages.

Ans: The parameter of interest here is proportion 

e) (1 pt) In a sample of one hundred recent college graduates, it is found that 85 percent expect to get a job within one year of their graduation date.

Ans: The parameter of interest here is proportion 

## 2

(5 pt) In 2013, the Pew Research Foundation reported that “45% of U.S. adults report that they live with one or more chronic conditions”. However, this value was based on a sample, so it may not be a perfect estimate for the population parameter of interest on its own. The study reported a standard error of about 1.2%, and a normal model may reasonably be used in this setting. Create a 95% confidence interval for the proportion of U.S. adults who live with one or more chronic conditions. Also interpret the confidence interval in the context of the study.



## 5 

An investigator is interested in understanding the relationship, if any, between the analytical
skills of young gifted children and the father’s IQ, the mother’s IQ, and hours of educational TV.
The data are here:

```{r 5}
library(openintro)
data(gifted)
```

```{r 5.a}
lm
```